### 2025-04-09 correlation.qmd
20250409 - psych package

First, let's examine descriptive statistics (height is measured in inches):

@@

A partial correlation examines the association between variables after controlling for one or more variables.
Below, we examine a partial [Pearson correlation](#sec-correlationExamplesPearson) between height and fantasy points after controlling for age.

@@

The [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] contains the `cor.table()` function that generates a correlation matrix of variables, including the $r$ value, number of observations ($n$), and $p$-value for each association.
Here is a Pearson correlation matrix:

@@

Here is a correlation matrix of variables that includes just the $r$ value and asterisks denoting the statistical significance of the association, for greater concision: $^\dagger p < .1$; $^* p < .05$; $^{**} p < .01$; $^{***} p < .001$.

@@

Here is a partial correlation matrix, controlling for the player's age:

>>>>

First, let's examine [descriptive statistics](#sec-descriptiveStatistics) of height and fantasy points (height is measured in inches):

@@

A partial correlation examines the association between variables after controlling for one or more variables.
Below, we examine a partial [Pearson correlation](#sec-correlationExamplesPearson) between height and fantasy points after controlling for age using the the `partial.r()` function from the `psych` package [@R-psych].

@@

The [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] contains the `cor.table()` function that generates a correlation matrix of variables, including the $r$ value, number of observations ($n$), asterisks denoting statistical significance, and $p$-value for each association.
The asterisks follow the following traditional convention for statistical significance: $^\dagger p < .1$; $^* p < .05$; $^{**} p < .01$; $^{***} p < .001$.
Here is a Pearson correlation matrix:

@@

Here is a correlation matrix of variables that includes just the $r$ value and asterisks denoting the statistical significance of the association, for greater concision.

@@

The [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] contains the `partialcor.table()` function that generates a partial correlation matrix.
Here is a partial correlation matrix, controlling for the player's age:
